---
title: "Montyhall Package Inspection"
output:
  html_document:
    theme: readable
    df_print: paged
    highlight: tango
    toc: true
params:
  name:
    value: x
---

```{r setup, include=FALSE}
knitr::opts_chunk$set( echo=FALSE, eval=TRUE, comment="" )
```


```{r}
github.name <- params$name
if( github.name == "x" )
{ github.name <- "" }  # add valid name to knit rmd for testing purposes 

url.1 <- "https://github.com/"
username <- github.name
url.2 <- "/montyhall/archive/refs/heads/main.zip"
URL <- paste0( url.1, username, url.2 )
```


```{r}
library( tools )
# underlining titles breaks text version: 
Rd2txt_options( underline_titles=FALSE )

wd       <- getwd()
folder   <- "PKG-FILES"
pkg.path <- paste0( wd, "/", folder )


# FUNCTIONS TO INSPECT 
rds <- c( "create_game.Rd", "select_door.Rd", "open_goat_door.Rd",
          "change_door.Rd", "determine_winner.Rd", "play_game.Rd", 
          "play_n_games.Rd" )

fxs <- gsub( ".Rd", "", rds )


# REMOVE PRIOR INSTALLATIONS, IF ANY
# THEN INSTALL VIA GITHUB 

install_montyhall <- function( username ) {
  
  options( show.error.messages=FALSE )
  try( detach( package:montyhall ) )
  remove.packages( pkgs="montyhall" )

  options( show.error.messages=TRUE )
  path <- paste0( username, "/montyhall" )
  devtools::install_github( path )
}



# PRINT CONTENTS OF THE DESCRIPTION FILE
# IN THE STUDENT PACKAGE 

print_description <- function() {
  use.these <- 
    c( "Author",  "Title",
       "Package", "Version",
       "Imports", "Depends",
       "Description", "Built" )  
  x <- read.dcf(
    file   = system.file( "DESCRIPTION", package = "montyhall" ),
    fields = use.these )
  write.dcf( x )
}



# DOWNLOAD ZIPPED PACKAGE FILES 
# FROM GITHUB AND SAVE LOCAL COPY 

download_package <- function( username, folder="PKG-FILES", URL=URL ) {

  if (! file.exists( folder ) )
  { dir.create( folder ) }
  setwd( folder )
  
  fn  <- paste0( "montyhall-", username, ".zip" )
  download.file( url=URL, destfile=fn )
    
  unzip( zipfile=fn )
}

# download_package( username="kphart2" )



# PRINT CONTENTS OF THE HELP FILES 

print_help <- function( fx="create_game.Rd" ) {
  
  tools::Rd2txt_options( underline_titles=FALSE )
  
  help.dir <- paste0( wd, "/", folder, "/montyhall-main/man" )
  setwd( help.dir )
  
  fx.name <- gsub( "_|\\.Rd", " ", fx ) |> toupper()
  
  cat( "#####   \n" )
  cat( paste0( "#####     ", fx.name, "\n" ) )
  cat( "#####   \n\n\n" )
  
  
  tools::Rd2txt( 
    Rd  = fx, 
    out = stdout(),
    package = "montyhall" )
  cat( "\n\n\n\n" )
  cat( "#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#\n\n\n\n")
}

# purrr::walk( rds, print_help )


# DELETE LOCAL UNZIPPED FILES
# IN THE MONTY DIRECTORY 

remove_pkg_files <- function() {
  setwd( pkg.path )
  unlink( "montyhall-main", recursive=TRUE, force=TRUE )
}


#  DELETE THE montyhall PACKAGE FOLDER 
#  IN THE LOCAL R INSTALLED PACKAGES DIRECTORY
#  TO ALLOW FOR A FRESH INSTALL AND 
#  PREVENT FILE CORRUPTION ERRORS
#  THAT APPEAR WHILE RUNNING EXAMPLES 

delete_pkg <- function() {
  lib.path <- (.libPaths())[1]
  pkg.path <- paste0( lib.path, "/montyhall" )
  unlink( pkg.path, recursive=TRUE, force=TRUE )
}


#  EXECUTE ALL TEST CODE PROVIDED 
#  IN THE @EXAMPLE SECTIONS OF THE
#  PACKAGE DOCUMENTATION 

run_example <- function( fx="create_game" ) {
  
  cat( "#####  \n" )
  cat( paste0("#####   ", fx, "\n" ) )
  cat( "#####  \n" )
  
  options( show.error.messages = FALSE )
  
  res <- try(
    
    utils::example( 
      topic          = fx, 
      package        = "montyhall", 
      lib.loc        = NULL, 
      character.only = TRUE,
      prompt.prefix  = "" ) )
  
  msg <- paste0( "\n>>>\n>>>  ", geterrmessage(), ">>>\n" )

  if( inherits( res, "try-error") )
  { cat( msg ) } 
  
  # else
  # { print( res$value ) }

  cat( "\n\n\n" )
  cat( "#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#\n\n\n\n\n")
}

# run_example( "create_game" )
```





## Install Package

<hr>

Installing package from:     **`r paste0( github.name, "/montyhall" )`** 

> [`r URL`](`r URL')

<hr>
<br>


```{r, echo=FALSE}
path <- paste0( github.name, "/montyhall" )
cat( paste0( "devtools::install_github( '", path,  "' )" ))
```

```{r, results="hide", warning=F, message=F}
install_montyhall( username=github.name )
```


```{r, warning=F, message=F }
library( montyhall )
```


## Description File 

```{r}
print_description()
```


## Inspect Documentation

```{r}
download_package( username=github.name )
purrr::walk( rds, print_help )
remove_pkg_files()
```

```{r, eval=F, echo=F}
# make sure package files deleted
print( wd )
print( pkg.path )
setwd( pkg.path )
dir()
fpath <- paste0( paste0( pkg.path, "/", "montyhall-main") )
unlink( fpath, recursive=TRUE, force=TRUE )
setwd( pkg.path )
dir()
```


## Run Examples

```{r}
purrr::walk( fxs, run_example )
```




```{r}
delete_pkg()
```






<style>

td{ padding: 6px 10px 6px 10px } 
th{ text-align: left; } 

h1{ font-size:calc(4em + 0.25vw) }
h2, h3, h4 { color: #995c00; }

h1 { margin-top:40px; }
h2 { margin-top:100px; }
h1 { margin-bottom:60px; }
h2 { margin-bottom:40px; }

body{
     font-family:system-ui,-apple-system,"Segoe UI",Roboto,Helvetica,Arial,sans-serif;
     font-size:calc(1.5em + 0.25vw);
     font-weight:300;
     line-height:1.65;
 }
 
pre{
     padding: 10px 10px 10px 60px;
}

</style>



