---
title: "Montyhall Package Inspection"
output:
  html_document:
    theme: readable
    df_print: paged
    highlight: tango
    toc: true
params:
  name: "x"
  pkg: "montyhall"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  eval = TRUE,
  comment = "",
  warning = FALSE,
  message = FALSE
)

options(width = 100)
tools::Rd2txt_options(underline_titles = FALSE)
```

```{r config}
github_name <- params$name
pkg_name  <- params$pkg 

root_dir  <- getwd()
files_dir <- file.path(root_dir, "PKG-FILES")
repo_path <- sprintf("%s/%s", github_name, pkg_name)

zip_url_main   <- sprintf("https://github.com/%s/%s/archive/refs/heads/main.zip", github_name, pkg_name)
zip_url_master <- sprintf("https://github.com/%s/%s/archive/refs/heads/master.zip", github_name, pkg_name)

dir.create(files_dir, showWarnings = FALSE, recursive = TRUE)

topics <- c(
  "create_game",
  "select_door",
  "open_goat_door",
  "change_door",
  "determine_winner",
  "play_game",
  "play_n_games"
)

rd_files <- paste0(topics, ".Rd")
```

```{r helpers}
safe_detach_package <- function(pkg) {
  pkg_search_name <- paste0("package:", pkg)
  if (pkg_search_name %in% search()) {
    try(detach(pkg_search_name, unload = TRUE, character.only = TRUE), silent = TRUE)
  }
}

safe_unload_namespace <- function(pkg) {
  if (pkg %in% loadedNamespaces()) {
    try(unloadNamespace(pkg), silent = TRUE)
  }
}

delete_installed_pkg <- function(pkg) {
  safe_detach_package(pkg)
  safe_unload_namespace(pkg)

  if (pkg %in% rownames(installed.packages())) {
    try(remove.packages(pkg), silent = TRUE)
  }

  invisible(TRUE)
}

install_montyhall <- function(username, pkg = "montyhall") {
  delete_installed_pkg(pkg)

  repo <- sprintf("%s/%s", username, pkg)
  devtools::install_github(
    repo = repo,
    upgrade = "never",
    dependencies = FALSE,
    quiet = TRUE
  )

  invisible(TRUE)
}

download_package <- function(username, folder, pkg = "montyhall") {
  dir.create(folder, showWarnings = FALSE, recursive = TRUE)

  urls <- c(
    main   = sprintf("https://github.com/%s/%s/archive/refs/heads/main.zip", username, pkg),
    master = sprintf("https://github.com/%s/%s/archive/refs/heads/master.zip", username, pkg)
  )

  for (branch in names(urls)) {
    this_url <- urls[[branch]]

    if (!httr::http_error(this_url)) {
      zip_file <- file.path(folder, sprintf("%s-%s-%s.zip", pkg, username, branch))
      ex_dir   <- file.path(folder, paste0(pkg, "-", branch))

      if (file.exists(zip_file)) file.remove(zip_file)
      if (dir.exists(ex_dir)) unlink(ex_dir, recursive = TRUE, force = TRUE)

      utils::download.file(this_url, destfile = zip_file, mode = "wb", quiet = TRUE)
      utils::unzip(zipfile = zip_file, exdir = folder)

      return(
        list(
          branch = branch,
          zip_file = zip_file,
          extracted_dir = ex_dir,
          url = this_url
        )
      )
    }
  }

  stop("Could not download package zip from either main or master branch.")
}

print_description <- function(pkg = "montyhall") {
  fields <- c(
    "Package", "Title", "Version", "Author",
    "Imports", "Depends", "Description", "Built"
  )

  desc_path <- system.file("DESCRIPTION", package = pkg)
  if (!nzchar(desc_path)) {
    cat("DESCRIPTION file not found for installed package.\n")
    return(invisible(NULL))
  }

  desc <- read.dcf(desc_path, fields = fields)
  write.dcf(desc)
  invisible(desc)
}

print_help <- function(rd_file, extracted_dir, pkg = "montyhall") {
  man_dir <- file.path(extracted_dir, "man")
  rd_path <- file.path(man_dir, rd_file)

  title <- gsub("_|\\.Rd", " ", rd_file)
  title <- toupper(title)

  cat("#####\n")
  cat(sprintf("#####   %s\n", title))
  cat("#####\n\n")

  if (!file.exists(rd_path)) {
    cat(sprintf("Help file not found: %s\n\n", rd_file))
    cat("#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#\n\n")
    return(invisible(NULL))
  }

  tools::Rd2txt(rd_path, out = stdout(), package = pkg)
  cat("\n\n#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#\n\n")

  invisible(TRUE)
}

find_rd_for_topic <- function(topic, extracted_dir) {
  man_dir <- file.path(extracted_dir, "man")

  if (!dir.exists(man_dir)) {
    stop(sprintf("man directory not found: %s", man_dir))
  }

  rd_files <- list.files(man_dir, pattern = "\\.Rd$", full.names = TRUE)

  for (rd_path in rd_files) {
    txt <- tryCatch(readLines(rd_path, warn = FALSE), error = function(e) character())

    if (any(grepl(sprintf("\\\\alias\\{%s\\}", topic), txt)) ||
        any(grepl(sprintf("\\\\name\\{%s\\}", topic), txt))) {
      return(rd_path)
    }
  }

  NULL
}

extract_example_code <- function(topic, extracted_dir) {
  rd_path <- find_rd_for_topic(topic = topic, extracted_dir = extracted_dir)

  ex_file <- tempfile(fileext = ".R")
  rd_obj  <- tools::parse_Rd(rd_path)
  tools::Rd2ex(rd_obj, out = ex_file)

  ex_file
}

run_example_capture <- function(topic, extracted_dir, pkg = "montyhall") {
  ex_file <- extract_example_code(
    topic = topic,
    extracted_dir = extracted_dir
  )

  suppressPackageStartupMessages(
    library( package="montyhall", character.only = TRUE)
  )

  ex_env <- new.env(parent = .GlobalEnv)

  capture.output(
    source(
      ex_file,
      local = ex_env,
      echo = TRUE,
      max.deparse.length = Inf,
      keep.source = TRUE
    )
  )
}


run_example <- function(topic, extracted_dir, pkg = "montyhall",
                        max_lines = 40) {
  cat("#####\n")
  cat(sprintf("#####   %s\n", topic))
  cat("#####\n")

  res <- try({
    output <- run_example_capture(
      topic = topic,
      extracted_dir = extracted_dir,
      pkg = pkg
    )

    if (length(output) == 0) {
      cat("(no printed output)\n")
    } else {
      cat(paste(head(output, max_lines), collapse = "\n"))
      cat("\n")

      if (length(output) > max_lines) {
        cat(sprintf(
          "\n... output truncated after %s of %s lines ...\n",
          max_lines, length(output)
        ))
      }
    }

    invisible(output)
  }, silent = TRUE)

  if (inherits(res, "try-error")) {
    cat("\n>>> ERROR\n")
    cat(gsub("\\s+", " ", as.character(res)), "\n")
  }

  cat("\n#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#\n\n")
  invisible(res)
}

cleanup_downloads <- function(folder, pkg = "montyhall") {
  candidates <- file.path(folder, c(paste0(pkg, "-main"), paste0(pkg, "-master")))
  for (x in candidates) {
    if (dir.exists(x)) unlink(x, recursive = TRUE, force = TRUE)
  }
  invisible(TRUE)
}

has_alias <- function(rd_path, topic) {
  if (!file.exists(rd_path)) return(FALSE)
  txt <- tryCatch(readLines(rd_path, warn = FALSE), error = function(e) character())
  any(grepl(sprintf("\\\\alias\\{%s\\}", topic), txt))
}

check_description_fields <- function(pkg = "montyhall") {
  required_fields <- c("Package", "Title", "Version", "Author", "Description")
  desc_path <- system.file("DESCRIPTION", package = pkg)

  if (!nzchar(desc_path) || !file.exists(desc_path)) {
    return(data.frame(
      check = paste("DESCRIPTION field:", required_fields),
      status = "FAIL",
      details = "DESCRIPTION file not found",
      stringsAsFactors = FALSE
    ))
  }

  desc_all <- read.dcf(desc_path)
  has_fields <- required_fields %in% colnames(desc_all)

  data.frame(
    check = paste("DESCRIPTION field:", required_fields),
    status = ifelse(has_fields, "PASS", "FAIL"),
    details = ifelse(has_fields, "Present", "Missing"),
    stringsAsFactors = FALSE
  )
}

check_help_files <- function(extracted_dir, rd_files, topics) {
  man_dir <- file.path(extracted_dir, "man")

  file_checks <- data.frame(
    check = paste("Help file exists:", rd_files),
    status = ifelse(file.exists(file.path(man_dir, rd_files)), "PASS", "FAIL"),
    details = ifelse(file.exists(file.path(man_dir, rd_files)), "Found in man/", "Missing from man/"),
    stringsAsFactors = FALSE
  )

  alias_checks <- data.frame(
    check = paste("Help alias exists:", topics),
    status = vapply(
      topics,
      function(topic) if (!is.null(find_rd_for_topic(topic, extracted_dir))) "PASS" else "FAIL",
      character(1)
    ),
    details = vapply(
      topics,
      function(topic) {
        hit <- find_rd_for_topic(topic, extracted_dir)
        if (is.null(hit)) "Matching alias missing" else paste("Found in", basename(hit))
      },
      character(1)
    ),
    stringsAsFactors = FALSE
  )

  rbind(file_checks, alias_checks)
}

check_examples <- function(topics, extracted_dir, pkg = "montyhall",
                           max_lines = 25) {

  results <- lapply(topics, function(topic) {
    res <- try({
      output <- run_example_capture(
        topic = topic,
        extracted_dir = extracted_dir,
        pkg = pkg
      )

      n_lines <- length(output) - 7
      if( n_lines < 0 ){ n_lines <- 0 }
      
      status <- if (n_lines > max_lines) "WARN" else "PASS"
      details <- if (n_lines > max_lines) {
        sprintf("Example produced %s lines of output (limit = %s)", n_lines, max_lines)
      } else {
        sprintf("Example completed (%s lines of output)", n_lines)
      }

      data.frame(
        check = paste("Example runs:", topic),
        status = status,
        details = details,
        stringsAsFactors = FALSE
      )
    }, silent = TRUE)

    if (inherits(res, "try-error")) {
      data.frame(
        check = paste("Example runs:", topic),
        status = "FAIL",
        details = gsub("\\s+", " ", as.character(res)),
        stringsAsFactors = FALSE
      )
    } else {
      res
    }
  })

  do.call(rbind, results)
}

check_installed_examples <- function(topics, pkg = "montyhall") {
  results <- lapply(topics, function(topic) {
    res <- try(
      utils::example(
        topic = topic,
        package = pkg,
        character.only = TRUE,
        prompt.prefix = ""
      ),
      silent = TRUE
    )

    if (inherits(res, "try-error")) {
      detail <- as.character(res)
      detail <- gsub("\n", " ", as.character(res))
      detail <- trimws(detail)

      status <- if (grepl("lazy-load database.*is corrupt", detail)) {
        "WARN"
      } else {
        "FAIL"
      }

      data.frame(
        check = paste("Installed help example:", topic),
        status = status,
        details = detail,
        stringsAsFactors = FALSE
      )
    } else {
      data.frame(
        check = paste("Installed help example:", topic),
        status = "PASS",
        details = "utils::example() completed without error",
        stringsAsFactors = FALSE
      )
    }
  })

  do.call(rbind, results)
}

build_grading_checks <- function(pkg, extracted_dir, topics, rd_files) {
  out <- rbind(
    data.frame(
      check = "Package installs",
      status = ifelse(pkg %in% rownames(installed.packages()), "PASS", "FAIL"),
      details = ifelse(pkg %in% rownames(installed.packages()), "Installed successfully", "Package not installed"),
      stringsAsFactors = FALSE
    ),
    data.frame(
      check = "Package loads",
      status = ifelse(paste0("package:", pkg) %in% search(), "PASS", "FAIL"),
      details = ifelse(paste0("package:", pkg) %in% search(), "Loaded successfully", "Package not attached"),
      stringsAsFactors = FALSE
    ),
    check_description_fields(pkg = pkg),
    check_help_files(extracted_dir = extracted_dir, rd_files = rd_files, topics = topics),
    check_examples(topics = topics, extracted_dir = extracted_dir, pkg = pkg)
  )

  out$status <- factor(out$status, levels = c("FAIL", "WARN", "PASS"))
  out[order(out$status, out$check), , drop = FALSE]
}
```

## Install Package

<hr>

Installing package from: **`r repo_path`**

```{r install-call, echo=FALSE}
cat(sprintf("devtools::install_github('%s')", repo_path))
```

```{r download-install}
download_info <- download_package(
  username = github_name,
  folder   = files_dir,
  pkg      = pkg_name
)

library(devtools)
install_montyhall(username = github_name, pkg = pkg_name)
library(package="montyhall", character.only = TRUE)
```

Zipped files downloaded from:

> [`r download_info$url`](`r download_info$url`)

<hr>
<br>

## Grading Checks

```{r grading-checks, results='hide'}
grading_checks <- build_grading_checks(
  pkg = pkg_name,
  extracted_dir = download_info$extracted_dir,
  topics = topics,
  rd_files = rd_files
)
```

### Summary

```{r grading-summary, results="asis"}
total_checks <- nrow(grading_checks)
n_pass <- sum(grading_checks$status == "PASS")
n_fail <- sum(grading_checks$status == "FAIL")

cat(sprintf("**Passed:** %s / %s  \n", n_pass, total_checks))
cat(sprintf("**Failed:** %s / %s", n_fail, total_checks))
```

### Checks Performed

```{r}
rownames(grading_checks) <- NULL

knitr::kable(
  grading_checks,
  align = c("l", "c", "l"),
  row.names = FALSE
)
```


## Description File

```{r description}
print_description(pkg = pkg_name)
```

## Inspect Documentation

```{r docs}
purrr::walk(
  rd_files,
  print_help,
  extracted_dir = download_info$extracted_dir,
  pkg = pkg_name
)
```


## Run Examples from Source .Rd Files

```{r run-examples}
purrr::walk(
  topics,
  run_example,
  extracted_dir = download_info$extracted_dir,
  pkg = pkg_name
)
```





```{r cleanup}
safe_detach_package(pkg_name)
safe_unload_namespace(pkg_name)
delete_installed_pkg(pkg_name)
cleanup_downloads(files_dir, pkg = pkg_name)
```



<style>

td{ padding: 6px 10px 6px 10px } 
th{ text-align: left; } 

h1{ font-size:calc(4em + 0.25vw) }
h2, h3, h4 { color: #995c00; }

h1 { margin-top:40px; }
h2 { margin-top:100px; }
h1 { margin-bottom:60px; }
h2 { margin-bottom:40px; }

body{
     font-family:system-ui,-apple-system,"Segoe UI",Roboto,Helvetica,Arial,sans-serif;
     font-size:calc(1.5em + 0.25vw);
     font-weight:300;
     line-height:1.65;
 }
 
pre{
     padding: 10px 10px 10px 60px;
}

</style>



