---
title: "Montyhall Package Inspection"
output:
  html_document:
    theme: readable
    df_print: paged
    highlight: tango
    toc: true
params:
  name:
    value: x
---

```{r setup, include=FALSE}
knitr::opts_chunk$set( echo=FALSE, eval=TRUE, comment="" )
```


```{r}
github.name <- params$name
if( github.name == "x" )
{ github.name <- "bmwaite" }  # add valid name to knit rmd for testing purposes 

url.1 <- "https://github.com/"
username <- github.name
url.2 <- "/montyhall/archive/refs/heads/main.zip"
URL <- paste0( url.1, username, url.2 )
```


```{r}
library( tools )
# underlining titles breaks text version: 
Rd2txt_options( underline_titles=FALSE )

wd       <- getwd()
folder   <- "PKG-FILES"
pkg.path <- paste0( wd, "/", folder )


# REMOVE PRIOR INSTALLATIONS, IF ANY
# THEN INSTALL VIA GITHUB 

install_montyhall <- function( username ) {

  pkg.location <- system.file( package="montyhall" )
  if( pkg.location == "" )
  {
    options( show.error.messages=FALSE )
    try( detach( package:montyhall ) )
    try( remove.packages( pkgs="montyhall" ) )    
  }
  
  options( show.error.messages=TRUE )
  path <- paste0( username, "/montyhall" )
  devtools::install_github( path )
}



# PRINT CONTENTS OF THE DESCRIPTION FILE
# IN THE STUDENT PACKAGE 

print_description <- function() {
  use.these <- 
    c( "Author",  "Title",
       "Package", "Version",
       "Imports", "Depends",
       "Description", "Built" )  
  x <- read.dcf(
    file   = system.file( "DESCRIPTION", package = "montyhall" ),
    fields = use.these )
  write.dcf( x )
}



# DOWNLOAD ZIPPED PACKAGE FILES 
# FROM GITHUB AND SAVE LOCAL COPY 

download_package <- function( username, folder="PKG-FILES", pkg.url ) {

  if (! file.exists( folder ) )
  { dir.create( folder ) }
  setwd( folder )
  
  fn  <- paste0( "montyhall-", username, ".zip" )
  download.file( pkg.url, destfile=fn )
    
  unzip( zipfile=fn )
}

# download_package( username="kphart2" )



# PRINT CONTENTS OF THE HELP FILES 

print_help <- function( fx="create_game.Rd" ) {
  
  tools::Rd2txt_options( underline_titles=FALSE )
  
  help.dir <- paste0( wd, "/", folder, "/montyhall-main/man" )
  setwd( help.dir )
  
  fx.name <- gsub( "_|\\.Rd", " ", fx ) |> toupper()
  
  cat( "#####   \n" )
  cat( paste0( "#####     ", fx.name, "\n" ) )
  cat( "#####   \n\n\n" )
  
  
  tools::Rd2txt( 
    Rd  = fx, 
    out = stdout(),
    package = "montyhall" )
  cat( "\n\n\n\n" )
  cat( "#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#\n\n\n\n")
}

# purrr::walk( rds, print_help )


# DELETE LOCAL UNZIPPED FILES
# IN THE MONTY DIRECTORY 

remove_unzipped_files <- function() {
  setwd( pkg.path )
  unlink( "montyhall-main", recursive=TRUE, force=TRUE )
}


#  DELETE THE montyhall PACKAGE FOLDER 
#  IN THE LOCAL R INSTALLED PACKAGES DIRECTORY
#  TO ALLOW FOR A FRESH INSTALL AND 
#  PREVENT FILE CORRUPTION ERRORS
#  THAT APPEAR WHILE RUNNING EXAMPLES 

delete_pkg <- function( pkgname ) {
  lib.path <- (.libPaths())[1]
  pkg.path <- paste0( lib.path, "/", pkgname )
  unlink( pkg.path, recursive=TRUE, force=TRUE )
}


#  EXECUTE ALL TEST CODE PROVIDED 
#  IN THE @EXAMPLE SECTIONS OF THE
#  PACKAGE DOCUMENTATION 

run_example <- function( fx="create_game" ) {
  
  cat( "#####  \n" )
  cat( paste0("#####   ", fx, "\n" ) )
  cat( "#####  \n" )
  
  options( show.error.messages = FALSE )
  
  res <- try(
    
    utils::example( 
      topic          = fx, 
      package        = "montyhall", 
      lib.loc        = NULL, 
      character.only = TRUE,
      prompt.prefix  = "" ) )
  
  msg <- paste0( "\n>>>\n>>>  ", geterrmessage(), ">>>\n" )

  if( inherits( res, "try-error") )
  { cat( msg ) } 
  
  # else
  # { print( res$value ) }

  cat( "\n\n\n" )
  cat( "#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~#\n\n\n\n\n")
}

# run_example( "create_game" )
```





```{r}
# MAKE SURE PACKAGE IS FRESH 
delete_pkg( "montyhall" )
```



## Install Package

<hr>

Installing package from:     **`r paste0( github.name, "/montyhall" )`** 

```{r, echo=FALSE}
path <- paste0( github.name, "/montyhall" )
cat( paste0( "devtools::install_github( '", path,  "' )" ))
```

Zipped files downloaded from: 

> [`r URL`](`r URL`)

<hr>
<br>



```{r, results="hide", warning=F, message=F}
download_package( username=github.name, pkg.url=URL )
install_montyhall( username=github.name )
library( montyhall )
```





## Run Examples


```{r, eval=F, echo=F}
# TEST CODE FROM EXAMPLE 

utils::example( 
  topic = "create_game", 
  package = "montyhall", 
  lib.loc = NULL, 
  character.only = TRUE,
  prompt.prefix = "" ) 
```


```{r}
# TEST ALL 

fxs <- 
  c("create_game", "select_door", "open_goat_door", "change_door", 
    "determine_winner", "play_game", "play_n_games")

purrr::walk( fxs, run_example )
```




## Description File 

```{r}
print_description()
```





## Inspect Documentation

```{r}

rds <- 
  c( "create_game.Rd", "select_door.Rd", "open_goat_door.Rd",
     "change_door.Rd", "determine_winner.Rd", "play_game.Rd", 
     "play_n_games.Rd" )


purrr::walk( rds, print_help )
```








```{r, warning=F, message=F}
# DELETE PACKAGE INSTALLATION
# FROM LOCAL R PKG LIBRARY 

detach( "package:montyhall", unload=TRUE )
remove.packages( "montyhall" )
unloadNamespace("montyhall")
delete_pkg( "montyhall" )
remove_unzipped_files()
```





<style>

td{ padding: 6px 10px 6px 10px } 
th{ text-align: left; } 

h1{ font-size:calc(4em + 0.25vw) }
h2, h3, h4 { color: #995c00; }

h1 { margin-top:40px; }
h2 { margin-top:100px; }
h1 { margin-bottom:60px; }
h2 { margin-bottom:40px; }

body{
     font-family:system-ui,-apple-system,"Segoe UI",Roboto,Helvetica,Arial,sans-serif;
     font-size:calc(1.5em + 0.25vw);
     font-weight:300;
     line-height:1.65;
 }
 
pre{
     padding: 10px 10px 10px 60px;
}

</style>



